---
title: "Problem set 1"
subtitle: "Hyoungchul Kim"
date: last-modified
format:
    pdf: default
    html: 
      self-contained: true
execute: 
  echo: fenced
reference-location: margin
citation-location: margin
---

# Basic and matching theory

## 1.

Let $h(X)$ be a function of $X$. Then expectation of mean squared error is: $\mathbb{E}\left( Y - h(X) \right)^2 = \mathbb{E} (Y - \mathbb{E}(Y|X) + \mathbb{E}(Y|X) + h(X))^2 = \mathbb{E}(Y-\mathbb{E}(Y|X))^2 + \mathbb{E}(\mathbb{E}(Y|X)+h(X))^2 - 2 \mathbb{E}\left[ (Y - \mathbb{E}(Y|X)) \underbrace{(\mathbb{E}(Y|X) - h(X))}_{\text{g(X)}} \right]$.

Then we can show that the last term becomes zero because $\mathbb{E}\left[ g(X)Y - g(X) \mathbb{E}(Y|X)\right] = \mathbb{E} \left[ g(X)Y - \mathbb{E}(g(X)Y|X) \right]= \mathbb{E}(g(X)Y) - \mathbb{E}(g(X)Y)$, which is due to law of iterated expectation. 

Then in order to minimize this error, we need to set $h(X)=\mathbb{E}(Y|X)$.

## 2. In potential outcomes framework, we can write ATT as $\mathbb{E}(\hat{\tau}_{ATT}) = \mathbb{E}\left[ Y_1 \mid D_i =1\right] - \mathbb{E}\left[ Y_0 \mid D_i =1\right] = \mathbb{E}\left[ Y_1 \mid D_i =1\right] - \mathbb{E}\left[ Y_i \mid D_i =1\right]$


# Matching empirics

## 1. 
 
Read the data and do basic data cleaning

First do it for whole sample
```{r}
#| code-fold: true
#| message: false
#| warning: false 
library(tidyverse)
library(modelsummary)
nsw = haven::read_dta("data/nsw.dta")
nsw_dw = haven::read_dta("data/nsw_dw.dta")
cps1 = haven::read_dta("data/cps1.dta")
psid1 = haven::read_dta("data/psid1.dta") 

# add up by data_id

data = nsw |>
  bind_rows(nsw_dw) |>
  bind_rows(cps1) |>
  bind_rows(psid1)

datasummary(age + education + black + hispanic + married + nodegree + re74 + re75 + re78 ~ data_id * (mean + sd), data = data)


```

This is for just treatment group

```{r}
#| cold-fold: true
#| message: false
#| warning: false 

data_treat = data |> filter(treat ==1)

datasummary(age + education + black + hispanic + married + nodegree + re74 + re75 + re78 ~ data_id *(mean + sd), data = data_treat)
```

